---
author: '[Nima Hejazi](https://nimahejazi.org) & Jeremy Coyle'
date: '`r format(Sys.time(), "%Y %b %d (%a), %H:%M:%S")`'
title: '[`sl3`](https://sl3.tlverse.org/): Simplifying machine learning in R
  using pipelines'
output:
  html_document:
    toc: yes
---

Based on materials originally produced by Jeremy Coyle, Nima Hejazi, Ivana
Malenica, and Oleg Sofrygin

Common in the language of modern data science are words such as "munging,"
"massaging," "mining" -- terms denoting the interactive process by which the
analyst extracts some form of deliverable inference from the data set at hand.
These terms express, among other things, the often convoluted process by which a
set of pre-processing and estimation procedures are applied to an input data set
in order to transform it into a
["tidy"](http://vita.had.co.nz/papers/tidy-data.html) data set from which
informative visualizations and summaries may be easily extracted. A formalism
that captures this involved process is that of machine learning _pipelines_. A
_pipeline_ -- popularized by the [method of the same
name](http://scikit-learn.org/stable/modules/generated/sklearn.pipeline.Pipeline.html)
in Python's [scikit-learn library](http://scikit-learn.org/stable/index.html) --
may be thought of as an _ordered_ set of instructions corresponding to
procedures to be applied to the input data set, with the ultimate goal of
producing a tidy output data set.

Recently, the new [`sl3` R package](https://github.com/tlverse/sl3) introduced
the pipeline idiom into the R programming language. A concrete understanding of
the utility of pipelines is best developed by example -- so, that's precisely
what we'll aim to do here. In the following, we'll apply the concept of a
machine learning pipeline to the canonical [iris data
set](https://en.wikipedia.org/wiki/Iris_flower_data_set), combining a series
of machine learning algorithms for classification with principal components
analysis, a simple pre-processing step for dimensionality reduction.

```{r setup_pkgs, message=FALSE}
library(datasets)
library(tidyverse)
library(data.table)
library(caret)
library(sl3)
set.seed(352)
```

In the above, we simply load a few of the R packages that we'll rely on
throughout this demonstration and set a seed in order to control any randomness
in the estimation procedure that follows. Next, let's load the iris data set:

```{r prepare_data, message=FALSE}
data(iris)
iris <- iris %>%
  as_tibble(.)
iris
```

As we see above, the iris data set consists of a simple structure: numerical
measurements of the length and width of sepals and petals, alongside the species
of the observed flower (restricted to three: _Iris setosa_, _Isis versicolor_,
_Iris virginica_).

To create very simple training and testing splits, we'll rely on the popular
[`caret` R package](https://topepo.github.io/caret/):

```{r split_data, message=FALSE}
trn_indx <- createDataPartition(iris$Species, p = .8, list = FALSE,
                                times = 1) %>%
  as.numeric()
tst_indx <- which(!(seq_len(nrow(iris)) %in% trn_indx))
```

Now that we have our training and testing splits, we can organize the data into
tasks -- the central bookkeeping object in the `sl3` framework. Essentially,
tasks represent a, well, data analytic _task_ that is to be solved by invoking
the various machine learning algorithms made available through `sl3`.

```{r make_iris_task}
# a task with the data from the training split
iris_task_train <- sl3_Task$new(
  data = iris[trn_indx, ],
  covariates = colnames(iris)[-5],
  outcome = colnames(iris)[5],
  outcome_type = "categorical"
)

# a task with the data from the testing split
iris_task_test <- sl3_Task$new(
  data = iris[tst_indx, ],
  covariates = colnames(iris)[-5],
  outcome = colnames(iris)[5],
  outcome_type = "categorical"
)

# let's take a look at the training data task
iris_task_train
```

Having set up the data properly, let's proceed to design _pipelines_ that we can
rely on for processing and analyzing the data. A __pipeline__ simply represents
a set of machine learning procedures to be invoked sequentially, with the
results derived from earlier algorithms in the pipeline being used to train
those later in the pipeline. Thus, a pipeline is a closed _end-to-end system_
for resolving the problem posed by an `sl3` task.

We'll rely on PCA for dimension reduction, gathering only the two most important
principal component dimensions to use in training our classification models.
Since this is a quick experiment with a well-studied data set, we'll use just
two classification procedures: (1) Logistic regression with regularization
(e.g., the LASSO) and (2) Random Forests.

```{r make_learners}
pca_learner <- Lrnr_pca$new(n_comp = 2)
glmnet_learner <- Lrnr_glmnet$new()
rf_learner <- Lrnr_randomForest$new()
```

Above, we merely instantiate the learners by invoking the `$new()` method of
each of the appropriate objects. We now have a machine learning object that
invokes PCA to generate and extract just the first two (from the argument
`n_comp` above) principal components derived from the design matrix.

Other than our PCA learner, we've also instantiated a regularized logistic
regression model (`glmnet_learner` above) based on the implementation available
through the popular [`glmnet` R
package](https://cran.r-project.org/package=glmnet), as well as a random forest
model based on the canonical implementation available in the
[`randomForest` R package](https://cran.r-project.org/package=randomForest).

Now that our individual learners are set up, we can intuitively string them into
pipelines by invoking the appropriate `$new()` method like so

```{r make_pipelines}
pca_to_glmnet <- Pipeline$new(pca_learner, glmnet_learner)
pca_to_rf <- Pipeline$new(pca_learner, rf_learner)
```

The first pipeline above merely invokes our PCA learner, extracting the first
two principal components of the design matrix from the input task and passing
these as inputs to the regularized logistic regression model. Similarly, the
second pipeline invokes PCA and passes the results to our random forest model.

To streamline the training of our pipelines, we'll bundle them into a single
_stack_, then train the model stack all at once. Similar in spirit to a
pipeline, a __stack__ is a bundle of `sl3` learner objects that are to be
trained together. The principle difference is that learners in a pipeline are
trained sequentially, as described above, while those in a stack are trained in
simultaneously. Thus, the models in a stack are trained independently of one
another.

Now, forward -- let's generate a stack and train the two pipelines on our
training split of the iris data set:

```{r train_stack}
model_stack <- Stack$new(pca_to_glmnet, pca_to_rf)
fit_model_stack <- model_stack$train(iris_task_train)
```

Having trained our stacked pipelines, we can now predict on the testing data set
simply by feeding the object `iris_task_test` that we created above to our model
stack using the `$predict()` method. After doing that, we'll simply do a bit of
bookkeeping to extract the predicted class probabilities (of each observation)
from the two pipelines in our stack.

```{r predict_and_cleanup}
out_model_stack <- fit_model_stack$predict(iris_task_test)
pipe_preds <- lapply(out_model_stack, unpack_predictions)
```

After extracting the predicted species probabilities for each observation (the
most likely iris species), we now clean up the results a bit, just to make them
easier to report

```{r get_class_preds}
# get class predictions
pipe1_classes <- predict_classes(pipe_preds[[1]])
pipe2_classes <- predict_classes(pipe_preds[[2]])
```

A standard way to summarize results in machine learning problems is the
[confusion
matrix](http://scikit-learn.org/stable/modules/generated/sklearn.metrics.confusion_matrix.html).
Now, let's take a look at the results, using the handy
[`confusionMatrix`
function](https://rdrr.io/cran/caret/man/confusionMatrix.html) from
[`caret`](https://topepo.github.io/caret/index.html) to compare our predicted
classes to the true species in the holdout/testing data set.

```{r confusion_mat_glmnet}
(cfmat_pipe1 <- confusionMatrix(pipe1_classes, iris_task_test$Y))
```

Let's find out whether our pipeline of PCA and Random Forest fared any better
than the one with PCA and GLMs above:

```{r confusion_mat_rf}
(cfmat_pipe2 <- confusionMatrix(pipe2_classes, iris_task_test$Y))
```

The predictions looks good!

---

## Summary

* We've taken a look at how to efficiently perform machine learning tasks with
  the `sl3` R package.
* We've examined standard machine learning idioms, including _tasks_,
  _pipelines_, and _stacks_.
* We've examined a simple case of training and predicting with pipelines on the
  canonical Iris data set.

